The Unicode Cookbook \linebreak for Linguists

### 2015-12-15 book/chapters/preface.tex
edits in preface

In the 1840s, several researchers and teachers began developing a notation to accurately transcribe articulatory gestures and their acoustic correlates. Their goal was not only transcription, they also wanted to use their notation for spelling reform (especially English), for teaching literacy and foreign languages, and in particular for developing writing systems for unwritten languages.

@@

The first instrumental measurements in the 1870-80s showed unequivocally that spoken language is comprised of sounds waves; it is dynamic and not discrete. Given the fluid nature of the speech stream, sounds differ from one utterance to the next and the same sound may differ in different environments. However, many of these changes are predictable in the context of those environments and they need not necessarily form distinctive lexical contrasts.

These findings left the early phoneticians in a conundrum given the aims of developing precise phonetic notion, but also practical alphabetic-based writing. It is in this vein that the phonetic and phonemic distinction was introduced very early into transcription. The etic vs emic principle took hold, in essence, even before terminology for the distinction was introduced and adopted.

A transcription system that showed both contrastive and non-contrastive elements emerged, taking form notably in the International Phonetic Alphabet (which itself has been updated from time-to-time over the last nearly 120 years). The association between phonetic and phonemic would be maintained in the International Phonetic Alphabet and through the various instantiations of phonological theory, which developed from the work of the early phoneticians.

The model for the principle of contrast was alphabetic writing, in particular shallow orthography, i.e.\ a one-to-one mapping between sounds and symbol. Traditional alphabetic orthography, however, did not scale to precise phonetic notation, so the former was sprinkled with diacritics. This dichotomy  causes many issues -- linguistic, theoretic, and recently technological -- which we bring to light in this book.

>>>>

In the 1840s, several researchers and teachers began developing notation to accurately transcribe articulatory gestures and their acoustic correlates. Their goal was not only transcription, they wanted to use their notation for spelling reform (especially English), for teaching literacy and foreign languages, and in particular for developing writing systems for unwritten languages.

@@

The first instrumental measurements in the 1870-80s showed unequivocally that spoken language is comprised of sounds waves; it is dynamic and not discrete. Given the fluid nature of the speech stream, sounds differ from one utterance to the next and the same sound may differ in different phonetic environments. However, many of these alterations are predictable in the context of their environments and they need not necessarily form distinctive lexical contrasts.

These findings left the early phoneticians in a conundrum given the aims of developing precise phonetic notion, but also practically-minded alphabetic-based writing. It is in this vein that the phonetic and phonemic distinction was introduced very early into transcription. The etic vs emic principle took hold, in essence, even before the dichotomy was widely accepted and before terminology for linguistic distinctiveness was introduced and adopted.

A transcription system that showed both contrastive and non-contrastive elements emerged, taking form notably in the International Phonetic Alphabet (which itself has been updated from time-to-time over the last nearly 120 years). The association between phonetic and phonemic would be maintained in the International Phonetic Alphabet and through the various instantiations of phonological theory, which developed on the work by early phoneticians.

The model for phonemic contrast was alphabetic writing, in particular shallow orthography, i.e.\ one-to-one mappings between sound and symbol. Traditional alphabetic orthography, however, did not scale to precise phonetic notation, so the former was sprinkled with diacritics to denote non-contrastive phenomena. This dichotomy between phonemic and phonetic and its implementation in writing and consistent encoding causes many issues -- linguistic, theoretic, and most recently technological -- which we bring to light in this book.
